%% uctest.tex 11/3/94
%% Copyright (C) 1988-2004 Daniel Gildea, BBF, Ethan Munson.
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License, either version 1.3
% of this license or (at your option) any later version.
% The latest version of this license is in
%   http://www.latex-project.org/lppl.txt
% and version 1.3 or later is part of all distributions of LaTeX
% version 2003/12/01 or later.
%
% This work has the LPPL maintenance status "maintained".
% 
% The Current Maintainer of this work is Daniel Gildea.
%
% 2007/08/01
% LaTeX Package "ucr" is modified from LaTeX package "ucthesis."
% This modification is therefore under to the conditions of 
% the LaTeX Project Public License.
% Its formality is suitable for the dissertation of Universty of
% California, Riverside.
% This test document is for the convenience of all students of
% Universty of California, Riverside.
% Contact Charles Yang at chcyang@yahoo.com if you like.
% Charles Yang has nothing to do with the original author's sarcasm.
%

% \documentclass[11pt]{ucthesis}
% \documentclass[11pt]{ucr}
\documentclass[oneside,final, letterpaper]{ucr}
\input{packages}

\graphicspath{{Pictures/}} % Specifies the directory where pictures are stored

\begin{document}
\dsp
% Declarations for Front Matter

\title{Measurement of Identified Charged Hadron Anisotropic Flow in d+Au Collisions}
\author{Matthew Benjamin Mendoza}
\degreemonth{December}
\degreeyear{2016}
\degree{Doctor of Philosophy}
\chair{Dr. Richard Seto}
\othermembers{Dr. Kenneth Barish\\
Dr. John Ellison}
\numberofmembers{3}
\field{Physics}
\campus{Riverside}

\maketitle
\copyrightpage{}
\approvalpage{}

\degreesemester{Fall}


\begin{acknowledgements}
The pursuit of my doctoral degree has lasted for nearly a decade of my life; almost a third of my years on Earth as of the moment I write these acknowledgments. I would like to express my gratitude to all those who were instrumental to my success through this time.

First and foremost, I am thankful to God for His faithfulness and providence through the ups and downs of life, both scholastic and personal.  

Deepest heartfelt gratitude to my advisor, Rich Seto, whose support, direction, and wisdom were essential and paramount to my success as a graduate student. I would also like to thank the other members of my dissertation committee, Ken Barish and John Ellison.

I am so very thankful for my wife, Katharine, for her patience and support through these many many years. Through moments when I felt as if I would never finish, she has consistently been there supporting and encouraging me all along the way.

Thanks to my parents, grandparents, and family for their love and support. I am eternally thankful for the direction, encouragement, and discipline that has brought me to where I am today. 

I would also like to thank the members of the UC Riverside heavy ion physics group for discussions, insight, encouragement, and help with various parts of this analysis and my graduate experience. Specifically, Richard and Aneta Hollis for their countless acts of assistance with many details of my analysis as well as their consistent availability to help me with problems that arose. To Mike Beaumier, Sky Rolnick, and Ken Sedgwick for help with complicated analysis code problems, and Tim Hester and David Kleinjan for various discussions on the subject matter as well as their friendship.

To John Babb, Jared Wong, and Arun Luthra for thirsty thursdays, scotch and cigar nights, and for being great sources of needed distraction during the many long days in the physics department.

To the men who have been great friends and pastors to me, Jonathan Bilima, Michael Stafford, and Avery Stafford.

To Jeremy Cunningham and Arthur Carpenter for being the best friends I could ever ask for.

To my undergraduate physics mentors, Alex Grillo, Jessica Metcalfe, Michael Dine, and Bruce Schumm for supporting and nurturing my passion for physics and research in such a formative time of my life. 

To my friends of \#foodchat for their camaraderie and support over these years. 
\end{acknowledgements}

\begin{dedication}
\null\vfil
{\large
\begin{center}
\textit{To my family for their unyielding support.}
\end{center}}
\vfil\null
\end{dedication}

\input{abstract}

\tableofcontents
\listoffigures
\listoftables
\end{frontmatter}

% \part{First Part}


% \part{First Part}

\input{Introduction} %usually intro
\input{NuclearMatter}
\input{Apparatus}
\input{HICollisions}
\input{Flow}
\input{EventPlane}
\input{Results}
\input{Systematics}
\input{Conclusions}
\input{appendix_coords}
\input{appendix_abbreviations}
\input{appendix_nuclmodfact}
\input{appendix_data}

\nocite{*}
%\singlespacing
%\bibliographystyle{alpha}
\ssp
\bibliographystyle{unsrt} 
\bibliography{references} 


\end{document}
